---
title: "Computational Modeling - Assignment 2"
author: "Riccardo Fusaroli"
date: "29/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. there is a second part at the bottom for next week.

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
library(brms)
library(rethinking)

d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

#Define Grid
dens=20
p_grid = seq(from=0, to = 1, length.out = dens)
#Define  flat Prior
prior_flat = rep(1,dens)

#check if it looks crazy
dens(rbinom(1e4, 9, runif(1e4, 0, 1))) 

#Compute likelihood
likelihood =dbinom(3, size =6, prob = p_grid)
plot(likelihood)

#Compute posterior
unstd.posterior=likelihood * prior_flat
posterior = unstd.posterior/sum(unstd.posterior)
plot(posterior)

#What is the probability he knows more than chance (0.5)?
sum(posterior[p_grid<0.5])

#plot
d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)
ggplot(d, aes(grid,posterior)) + geom_point() +geom_line()+theme_classic()+ geom_line(aes(grid, prior/dens),color= 'red')+ xlab("Knowledge of CogSci")+ ylab("posterior probability")

# Prøv med en Quadric approximation
library(rethinking)
globe.qa = map(
  alist(
    w~dbinom(6,p), #binomial
    p~dunif(0,1) #uniform
  ),
  data=list(w=3))
precis(globe.qa)
# Det som MAP funktionen viser er at ved at "bestige bjerget" til the "maximum a posteriori estimate" hvor posterior's peak vil være beliggende. = tilsvarenede "mean" value i precis outputtet. 

#_____ Riccardo's rmd kode.


# Model sampling only from the prior (for checking the predictions your prior leads to)
FlatModel_priorCheck <- brm(Correct|trials(Questions) ~ 1, 
                 data = subset(d, Teacher=="RF"),
                 prior = prior("uniform(0,1)", class = "Intercept"),
                 family = binomial,
                 sample_prior = "only") # here we tell the model to ignore the data

# Plotting the predictions of the model (prior only) against the actual data
pp_check(FlatModel_priorCheck, nsamples = 100)

# Model sampling by combining prior and likelihood
FlatModel <- brm(Correct|trials(Questions) ~ 1, 
                 data = subset(d, Teacher=="RF"),
                 prior = prior("uniform(0,1)", class = "Intercept"),
                 family = binomial,
                 sample_prior = T)
# Plotting the predictions of the model (prior + likelihood) against the actual data
pp_check(FlatModel, nsamples = 100)

# plotting the posteriors and the sampling process
plot(FlatModel)


PositiveModel_priorCheck <- brm(Correct|trials(Questions) ~ 1,
                     data = subset(d, Teacher=="RF"),
                     prior = prior("normal(0.8,0.2)", 
                                   class = "Intercept"),
                     family=binomial,
                     sample_prior = "only")
pp_check(PositiveModel_priorCheck, nsamples = 100)

PositiveModel <- brm(Correct|trials(Questions) ~ 1,
                     data = subset(d, Teacher=="RF"),
                     prior = prior("normal(0.8,0.2)", 
                                   class = "Intercept"),
                     family=binomial,
                     sample_prior = T)
pp_check(PositiveModel, nsamples = 100)
plot(PositiveModel)

SkepticalModel_priorCheck <- brm(Correct|trials(Questions) ~ 1, 
                      data = subset(d, Teacher=="RF"),
                      prior=prior("normal(0.5,0.01)", class = "Intercept"),
                      family=binomial,
                      sample_prior = "only")
pp_check(SkepticalModel_priorCheck, nsamples = 100)

SkepticalModel <- brm(Correct|trials(Questions) ~ 1, 
                      data = subset(d, Teacher=="RF"),
                      prior = prior("normal(0.5,0.01)", class = "Intercept"),
                      family = binomial,
                      sample_prior = T)
pp_check(SkepticalModel, nsamples = 100)
plot(SkepticalModel)




# kat

#Define Grid
dens=1000
p_grid = seq(from=0, to = 1, length.out = dens)
#Define  flat Prior
prior_flat = rep(1,dens)
#Compute likelihood
likelihood =dbinom(3, size =6, prob = p_grid)
#Compute posterior
unstd.posterior=likelihood * prior_flat
posterior = unstd.posterior/sum(unstd.posterior)
#What is the probability he knows more than chance (0.5)?
sum(posterior[p_grid<0.5])
#Riccardo's probability that he knows more than 50% is 50% (Might differ due to samples)
#Plot
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior_flat,likelihood=likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")


```



2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r}

#2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
#2a. Produce plots of the prior, and posterior for each teacher.


#Define Grid
dens=1000
p_grid = seq(from=0, to = 1, length.out = dens)
#Define Prior
prior_flat = rep(1,dens)


#RICCARDO
#Compute likelihood
likelihood_riccardo =dbinom(3, size =6, prob = p_grid)
#Compute posterior
unstd.posterior_riccardo=likelihood_riccardo * prior_flat
posterior_riccardo = unstd.posterior_riccardo/sum(unstd.posterior_riccardo)
#------------------------
#KRISTIAN
#Compute likelihood
likelihood_Kristian <- dbinom(2, size =2, prob = p_grid)
#Compute posterior
unstd.posterior_K=likelihood_Kristian * prior_flat
posterior_Kristian = unstd.posterior_K/sum(unstd.posterior)
#-------------------------------
#JOSH
#Compute likelihood
likelihood_Josh =dbinom(160, size =198, prob = p_grid)
#Compute posterior
unstd.posterior_J=likelihood_Josh * prior_flat
posterior_Josh = unstd.posterior_J/sum(unstd.posterior)
#-------------------------------------

#Compute likelihood
likelihood_mikkel =dbinom(66, size =132, prob = p_grid)
#Compute posterior
unstd.posterior_M=likelihood_mikkel * prior_flat
posterior_mikkel = unstd.posterior_M/sum(unstd.posterior)
posterior_mikkel



#Produce plots for all teachers
library(ggplot2)

install.packages(gridExtra)
library(gridExtra)

Data_riccardo=data.frame(grid=p_grid,posterior=posterior_riccardo,prior=prior_flat,likelihood=likelihood_riccardo)
p1 = ggplot(Data_riccardo,aes(grid,posterior_riccardo))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
p1

Data_Kristian=data.frame(grid=p_grid,posterior=posterior_Kristian,prior=prior_flat,likelihood=likelihood_Kristian)
p2 = ggplot(Data_Kristian,aes(grid,posterior_Kristian))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
p2

Data_Josh=data.frame(grid=p_grid,posterior=posterior_Josh,prior=prior_flat,likelihood=likelihood_Josh)
p3 = ggplot(Data_Josh,aes(grid,posterior_Josh))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
p3

Data_mikkel=data.frame(grid=p_grid,posterior=posterior_mikkel,prior=prior_flat,likelihood=likelihood_mikkel)
p4=ggplot(Data_mikkel,aes(grid,posterior_mikkel))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
p4

grid.arrange(p1, p2, p3, p4, nrow = 1)

#Who is the best?
#Riccardo
p_grid[which.max(posterior_riccardo)]
RiccardoSamples = sample(p_grid , prob=posterior_riccardo , size=1e5 , replace=TRUE )
dens(RiccardoSamples)
HPDI(RiccardoSamples, prob = 0.5)
#the highest posterior density interval for Riccardo is from 0.37 to 0.61  cog sci knowledge (50 % interval)
#Kristian
p_grid[which.max(posterior_Kristian)]
KristianSamples = sample(p_grid , prob=posterior_Kristian , size=1e5 , replace=TRUE )
dens(KristianSamples)
HPDI(KristianSamples, prob = 0.5)
#the highest posterior density interval for Krisitan is from 0.79 to 1  cog sci knowledge (50 % interval)
#Josh
p_grid[which.max(posterior_Josh)]
JoshSamples = sample(p_grid , prob=posterior_Josh, size=1e5 , replace=TRUE )
dens(JoshSamples)
HPDI(JoshSamples, prob = 0.5)
#the highest posterior density interval for Josh is from 0.78 to 0.82  cog sci knowledge (50 % interval)
#Mikkel
p_grid[which.max(posterior_mikkel)]
MikkelSamples = sample(p_grid, prob=posterior_mikkel , size=1e5 , replace=TRUE )
dens(MikkelSamples)
HPDI(MikkelSamples, prob = 0.5)
#the highest posterior density interval for Mikkel is from 0.47 to 0.52  cog sci knowledge (50 % interval)
#Kristian is "the best" since his MAP is 1
#Comment on mikkel and riccardo
#Riccardo and Mikkel both MAP values of 50 %, however there is much more uncertainty regarding Riccardo's estimate. The 50% most likely values for Riccardo's cog sci knowledge (HPDI) is between 0.36 and 0.60. For Mikkel, these values lie between 0.47 and 0.52.

```

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)

```{r}


#2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
#2a. Produce plots of the prior, and posterior for each teacher.

```{r}
#Define Grid
dens=1000
p_grid = seq(from=0, to = 1, length.out = dens)
#Define Prior
prior_flat_riccardo = rep(1,dens)
#Compute likelihood
likelihood_riccardo =dbinom(3, size =6, prob = p_grid)
#Compute posterior
unstd.posterior_riccardo=likelihood_riccardo * prior_flat_riccardo
posterior_riccardo = unstd.posterior_riccardo/sum(unstd.posterior_riccardo)
------------------------
#Define Grid
dens=1000
p_grid = seq(from=0, to = 1, length.out = dens)
#Define Prior
prior_flat_Kristian = rep(1,dens) 
#Compute likelihood
likelihood_Kristian =dbinom(2, size =2, prob = p_grid)
#Compute posterior
unstd.posterior=likelihood_Kristian * prior_flat_Kristian
posterior_Kristian = unstd.posterior/sum(unstd.posterior)
-------------------------------
#Define Grid
dens=1000
p_grid = seq(from=0, to = 1, length.out = dens)
#Define Prior
prior_flat_Josh = rep(1,dens)
#Compute likelihood
likelihood_Josh =dbinom(160, size =198, prob = p_grid)
#Compute posterior
unstd.posterior=likelihood_Josh * prior_flat_Josh
posterior_Josh = unstd.posterior/sum(unstd.posterior)
-------------------------------------
#Define Grid
dens = 1000
p_grid = seq(from=0, to = 1, length.out = dens)
#Define Prior
prior_flat_mikkel = rep(1,dens)
#Compute likelihood
likelihood_mikkel =dbinom(66, size =132, prob = p_grid)
#Compute posterior
unstd.posterior=likelihood_mikkel * prior_flat_mikkel
posterior_mikkel = unstd.posterior/sum(unstd.posterior)
#Produce plots for all teachers
library(ggplot2)
Data_riccardo=data.frame(grid=p_grid,posterior=posterior_riccardo,prior=prior_flat_riccardo,likelihood=likelihood_riccardo)
p1 = ggplot(Data_riccardo,aes(grid,posterior_riccardo))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat_riccardo/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
Data_Kristian=data.frame(grid=p_grid,posterior=posterior_Kristian,prior=prior_flat_Kristian,likelihood=likelihood_Kristian)
p2 = ggplot(Data_Kristian,aes(grid,posterior_Kristian))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat_Kristian/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
Data_Josh=data.frame(grid=p_grid,posterior=posterior_Josh,prior=prior_flat_Josh,likelihood=likelihood_Josh)
p3 = ggplot(Data_Josh,aes(grid,posterior_Josh))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat_Josh/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
Data_mikkel=data.frame(grid=p_grid,posterior=posterior_mikkel,prior=prior_flat_mikkel,likelihood=likelihood_mikkel)
p4=ggplot(Data_mikkel,aes(grid,posterior_mikkel))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior_flat_mikkel/dens),color='red')+  xlab("probability of correct")+ ylab("posterior probability")
#Plots for all teachers
ggarrange(p1, p2, p3,p4 + rremove("x.text"), 
          labels = c("Riccardo", "Kristian", "Josh","Mikkel"),
          ncol = 2, nrow = 2)
#Who is the best?
#Riccardo
p_grid[which.max(posterior_riccardo)]
RiccardoSamples = sample(p_grid , prob=posterior_riccardo , size=1e5 , replace=TRUE )
dens(RiccardoSamples)
HPDI(RiccardoSamples, prob = 0.5)
#the highest posterior density interval for Riccardo is from 0.37 to 0.61  cog sci knowledge (50 % interval)
#Kristian
p_grid[which.max(posterior_Kristian)]
KristianSamples = sample(p_grid , prob=posterior_Kristian , size=1e5 , replace=TRUE )
dens(KristianSamples)
HPDI(KristianSamples, prob = 0.5)
#the highest posterior density interval for Krisitan is from 0.79 to 1  cog sci knowledge (50 % interval)
#Josh
p_grid[which.max(posterior_Josh)]
JoshSamples = sample(p_grid , prob=posterior_Josh, size=1e5 , replace=TRUE )
dens(JoshSamples)
HPDI(JoshSamples, prob = 0.5)
#the highest posterior density interval for Josh is from 0.78 to 0.82  cog sci knowledge (50 % interval)
#Mikkel
p_grid[which.max(posterior_mikkel)]
MikkelSamples = sample(p_grid, prob=posterior_mikkel , size=1e5 , replace=TRUE )
dens(MikkelSamples)
HPDI(MikkelSamples, prob = 0.5)
#the highest posterior density interval for Mikkel is from 0.47 to 0.52  cog sci knowledge (50 % interval)
#Kristian is "the best" since his MAP is 1
#Comment on mikkel and riccardo
#Riccardo and Mikkel both MAP values of 50 %, however there is much more uncertainty regarding Riccardo's estimate. The 50% most likely values for Riccardo's cog sci knowledge (HPDI) is between 0.36 and 0.60. For Mikkel, these values lie between 0.47 and 0.52.
```
